---
title: "Statistical Inference Course Project - 1"
author: "Victor"
date: "7/20/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simulation Exercise

### Overview

In this project, we will investigate the exponential distribution in R and compare it with the Central Limit Theorem, illustrating via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials.

### Simulations

```{r libraries}
library(ggplot2)
```

We will simulate the exponential distribution with exp(n, lambda) where lambda is the rate parameter and n = 40. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda.

We have set lambda = 0.2 for all of the simulations. We will need a thousand simulations.

```{r simulation.exp}
lambda = .2
m0 = 1/lambda
sd0 = 1/lambda
n = 40
mns = NULL
for (i in 1 : 1000) mns = c(mns, mean(rexp(n, lambda)))
```

### Sample Mean versus Theoretical Mean

We will show the sample mean and compare it to the theoretical mean of the distribution.

First, let's see the sample mean.

```{r mean}
m1 <- mean(mns)
m1
m1 - m0
hist(mns, main = "Sample mean", xlab = "Sample mean")
```

If we compare the sample mean with the theoretical mean of the distribution (1/lambda = 5), we have observed that the distribution of means is centered in the theoretical mean.

### Sample Variance versus Theoretical Variance

Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

```{r variance}
v0 <- (sd0^2/n)
v1 <- var(mns)
v1 - v0
```

If we compare the sample variance with the theoretical variance of the distribution (sd squared divided by n), we have observed that they only differ in -0.01, so they can be compared.

### Distribution

If we apply the CLT, we have to calculate the z-score.

```{r normal.2}
z.score <- (mns - m0) / (sd0 / sqrt(n))
p2 <- ggplot(as.data.frame(z.score), aes(x = z.score))
p2 <- p2 + stat_function(fun = dnorm, color = "red")
p2 <- p2 + stat_function(fun = dnorm, args = list(mean = mean(z.score), sd = sd(z.score)), color = "blue")
p2 <- p2 + labs(title = "Comparision of the distribution with a normal distribution 2", x = "Means", y = "Density")
p2
```

The conclusion is the same. The distribution is centered at zero and it is really similar to the one of a normal distribution.

## Appendix material

We will show that the distribution is approximately normal comparing them with the density function (dnorm).

```{r normal.1}
p1 <- ggplot(as.data.frame(mns), aes(x = mns))
p1 <- p1 + stat_function(fun = dnorm, args = list(mean = m0, sd = sqrt(v0)), color = "red")
p1 <- p1 + stat_function(fun = dnorm, args = list(mean = m1, sd = sqrt(v1)), color = "blue")
p1 <- p1 + labs(title = "Comparision of the distribution with a normal distribution 1", x = "Means", y = "Density")
p1
```

As we can see, the distribution (blue) is approximately that of a standard normal (red).